Designing an effective intervention of motivational interview using sequence classification methods

### 2017-03-01 paper/AMIA/main.tex
update dataset statistics
update data collection and part of result

\section*{Methods}
\subsection*{\textit{Data collection}}
Table~\ref{tab:data_dist} shows the statistics of used data sets. The experimental data sets for this work were constructed based on the transcripts of 37 motivational interview sessions, which include a total of 11,353 segmented and annotated utterances. A fragment of an adolescent session transcript is presented in Table~\ref{tab:anno_examp}. The utterances were annotated based on MYSCOPE codebook \cite{carcone2013provider} with 115 different behavior codes, which are grouped into the youth, caregiver, and counselor code groups. These utterances are then scanned from first to last and divided into successful and unsuccessful sequences when a commitment or change talk is found. A total of 1072 sequences have been observed which further partitioned into two subsets based on the outcome of motivational interviews: one data set that includes all sequences of positive change talk or commitment language (880 sequences) and the other data set which includes utterances ended with negative change talk or commitment language (192 sequences). In our experiment, these two subsets are utilized to train Markov chain and HMM. \\

\begin{table}[h]
\centering
\caption{Dataset statistics.}
\label{tab:data_dist}
  \begin{tabular}{|l|l|l|l|l|l|}
  \hline
   \textbf{Sequence type} & \textbf{Total sequences}  & \textbf{Successful sequences}  & \textbf{Unsuccessful sequences} & \textbf{Avg. length} \\ \hline      
 Normal sequences & 1072 & 880 (82.09\%) & 192 (17.91\%) & 3.5 \\\hline
Alternate sequences & 18000 & 15000 (82.09\%) & 3000 (17.91\%) & 4.5 \\\hline 
  \end{tabular}
\end{table}

@@

\section*{Results}
Experimental evaluation of automatic annotation of clinical interview fragments and their sequences by using machine learning and probabilistic model included several dimensions:
\begin{itemize}
\item determining the performance of classifiers on the codebooks of different size;
\item determining the effectiveness of the proposed contextual and semantic features.
\item determining the performance of Markov model in conjunction with determining PPC sequences that are most likely to translate into CT.
\end{itemize}

\subsection*{\textit{Performance of models for the classification of normal PPC code sequences}}
Since clinical researchers tried to increase the desire and ability of adolescent to change their current behavior to target behavior, we used sequence of successful and unsuccessful codes only from adolescent session for our sequential analysis. Summary of performance for the classification of sequential data is presented in Table~\ref{table:sequence}. From Table~\ref{table:sequence}, it follows that our designed Markov model works very well in terms of precision and achieves precision 0.7574 with F1-measure 0.7092. The strength of the model for the classification of each type of sequences is illustrated by providing the confusion matrix in Table~\ref{table:conf}. It also shows that successful patterns are identified more accurately compare to unsuccessful motivational interview sequences due to the imbalance of data.\\

>>>>

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental data sets for this work were constructed based on the transcripts of 37 motivational interview sessions, which include a total of 11,353 segmented and annotated utterances. Since clinical researchers tried to increase the desire and ability of adolescent to change their current behavior to target behavior, we used sequences from adolescent session and excluded sequences from caregiver session in the MI transcript. The utterances were annotated based on MYSCOPE codebook \cite{carcone2013provider} with 115 different behavior codes, which are grouped into the youth, caregiver, and counselor code groups. These utterances are then scanned from first to last and divided into successful and unsuccessful sequences when a commitment or change talk is found. A total of 1072 sequences have been observed which further partitioned into two subsets based on the outcome of motivational interviews: one data set that includes all sequences of positive change talk or commitment language (880 sequences) and the other data set which includes utterances ended with negative change talk or commitment language (192 sequences). Alternate sequences are the possible combination of normal sequence. For example, there are 6 possible alternate sequences for the following sequences: (C1, A1, A2, C2, C3, C4, A3), where adolescent and counselor codes are represented by A and C, respectively. Therefore, (C1, A1, C2, A3), (C1, A1, C3, A3), (C1, A1, C4, A3), (C1, A2, C2, A3), (C1, A2, C3, A3) and (C1, A2, C4, A3) were used as alternate sequences for the above sequence. A total of 18753 alternate sequences have been observed which further partitioned into two subsets of 18139 successful sequences and 614 unsuccessful sequences. In our experiment, normal and alternate sequences are employed with two subsets (successful and unsuccessful) to train Markov chain and HMM. Dataset statistics and a fragment of an adolescent session transcript are presented in Table~\ref{tab:data_dist} and Table~\ref{tab:anno_examp}. \\

\begin{table}[h]
\centering
\caption{General statistics of dataset.}
\label{tab:data_dist}
  \begin{tabular}{|l|l|l|l|l|l|}
  \hline
   \textbf{Sequence type} & \textbf{Total sequences}  & \textbf{\# Successful sequences}  & \textbf{\# Unsuccessful sequences} & \textbf{Avg. length} \\ \hline      
 Normal sequences & 1072 & 880 (82.09\%) & 192 (17.91\%) & 5.83 \\\hline
Alternate sequences & 18753 & 18139 (96.72\%) & 614 (3.28\%) & 12.89 \\\hline 
  \end{tabular}
\end{table}

@@

\section*{Results}
Experimental evaluation of the classification of utterance sequences by using probabilistic models are reported with two set of experiments: first set of experiment was focused on normal sequences, while the second set of experiment was performed on alternate sequences.

\subsection*{\textit{Performance of models for the classification of normal PPC code sequences}}
Summary of performance for the classification of sequential data is presented in Table~\ref{table:sequence}. From Table~\ref{table:sequence}, it follows that our designed Markov model works very well in terms of precision and achieves precision 0.7574 with F1-measure 0.7092. The strength of the model for the classification of each type of sequences is illustrated by providing the confusion matrix in Table~\ref{table:conf}. It also shows that successful patterns are identified more accurately compare to unsuccessful motivational interview sequences due to the imbalance of data.\\
